---
title: "Untitled"
author: "Iwen Su"
date: "March 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)
library(corrplot)

```

### Read in data: column graph of Texas oil spills

```{r}

oil_spills <- read_csv("oil_spills.csv")

df <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` < 2017) %>% 
  group_by(`Accident Year`) %>% 
  summarise(Loss = sum(`Net Loss (Barrels)`))

colnames(df) <- c("Year","Loss")

ggplot(df, aes(x = Year, y = Loss)) +
  geom_col()

```

### Leaflet plot of spill locations in TX in 2016

```{r}

df_loc <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` == 2016) %>% 
  select(Latitude, Longitude, `Net Loss (Barrels)`)

colnames(df_loc) <- c("latitude", "longitude", "net_loss")

# Convert to simple features since R doesn't know this data frame is spatial data yet! But we have lat-lon data
class(df_loc)

# order of longitude and longitude columns must be as follows
oil_sf <- st_as_sf(df_loc, coords = c("longitude","latitude"), crs = 4326)
class(oil_sf)

leaflet(oil_sf) %>% 
  addTiles() %>% 
  addMarkers()

```

### Point Pattern Analysis

tmap plot with the Texas state shapefile

```{r}

# want to read in any layer that starts with prefix 'states' (e.g. states.shp, states.sbn...)
states <- st_read(dsn = ".", layer = "states")

tex_border <- states %>% 
  filter(STATE_NAME == "Texas") %>% 
  st_transform(4326) # transform to the coordinate you want
  
# st_crs(tex_border) - to check what the coordinate reference system is
  
plot(tex_border)

tm_shape(tex_border) +
  tm_polygons() +
  tm_shape(oil_sf) +
  tm_dots(size = 0.3)

  
```

### Convert the data to spatial points patterns (combination of point data and the bounding window)

Took spatial info and converted into something that R recognizes for point pattern analysis

```{r}

spill_sp <- as(oil_sf, "Spatial")
spill_ppp <- as(spill_sp, "ppp") # need to convert data to 'ppp' in order to do point pattern analysis

tx_sp <- as(tex_border, "Spatial") # create outer window for doing analysis
tw_owin <- as(tx_sp, "owin")

all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tw_owin) # corresponding longitude and longitude; bounding window is tw_owin

all_ppp 

```

### Create a density plot:


```{r}

plot(density(all_ppp, sigma = 0.4)) # radius is 0.4

```

### Quadrat test for spatial evenness

Are oil spills evenly distributed. Counts the number of events in each quadrant as well as the intensity and then use a chi-squared test to test for evenness.

```{r}

oil_qt <- quadrat.test(all_ppp, nx = 5, ny = 5) # 5 regions that exist horizonatally and vertically
oil_qt

# Reject - the data are not CSR (data is not evenly distributed); p value < 0.05

plot(all_ppp)
plot(oil_qt, add = TRUE, cex = 0.4)

# some quadrants are not equally sized, so the ppp analysis will approximate number of events if each block was the same size 


```

Evaluating whether data follows CSR (which follows a poisson distribution)
* If data more clustered than CSR? Then nearest neighbors will be closer together on average
* If data less clustered than CSR? Then nearest neighbors will be further apart on average

### G-Function for Nearest Neighbor Analysis

```{r}

lag <- seq(0,1, by = 0.01) # based on projection units (e.g. degree) - could find maximum distance between any nearest neighbor and a point to approximate where the data starts to peter (converge around 1.00 y-axis) 
oil_gfun <- envelope(all_ppp, fun = Gest, r = lag, nsim = 100)  # run 100 simulations for our CSR

# theo is our CSR and lo and hi confidence interval, obs is our actual observations

ggplot(oil_gfun, aes(x = r, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x = r, y = theo), color = "red")

# our observed data has a higher proportion of point pairs with nearest neighbors at shorter distances (closer together) compared to CSR data -- data is more clustered than CSR

```


### Nearest neighbor using the L-function (Ripley's K, standardized)

```{r}

r2 <- seq(0,3, by = 0.5) # chose a large increment for now just so it won't take a long time, realistically probably use a smaller increment

oil_lfun <- envelope(all_ppp, fun = Lest, r = r2, nsim = 20, global = TRUE) 
# looks at every single point in space that has an event/oil spill and makes increasing bubble around it until it incorporates all other obs that exist
# global = TRUE just makes sure you incorporate all your points

ggplot(oil_lfun, aes(x = r2, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x = r2, y = theo), color = "dodgerblue")

# Results: higher densities of neighbors that exist at each event (bubble) compared to CSR

```

**Conclusions: Considering nearest neighbors and global neighbors, we found that there is greater clustering in our data compared to CSR.**



